Elastic Scattering in General Relativistic Ray Tracing for Neutrinos

### 2018-02-21 paper.tex
Add conclusion

\section{Conclusions}
\label{sec:conclusions}
Add some conclusions.
\footnote{TODO (Brett): note some improvements needed: implement implicit time-stepping,
  use finite-difference grid for fluid vars, improve gray spectral assumptions}

>>>>

\section{Conclusions}
\label{sec:conclusions}
We have introduced a new general relativistic ray tracing method to compute
neutrino distribution functions around compact objects in dynamical configurations,
and which incorporates the effects of elastic scattering for the first time
within a ray tracing framework.
Elastic scattering of neutrinos into and out of each ray is included in our
method by using estimates of the background neutrino fields from an
M1 transport simulation.
To capture the energy spectrum of the background field, we have described a
\emph{spectral} method which uses neutrino energy densities over multiple
energy groups as input, and a
\emph{gray} method which uses neutrino energy and number densities averaged
over all energies as input.
We have also successfully tested the ray tracing code with a comprehensive
battery of tests.

In our tests (Sec.~\ref{sec:tests}) we have confirmed that elastic scattering
plays a significant role in redistributing neutrino energy- and
angle-distributions in common compact-object configurations.
The largest effects are seen in $\nu_x$ distributions
and to a lesser extent $\bar{\nu}_e$,
with the dominant effect a decrease in average energies from the central body,
and an increase in average energies from the scattering envelope.
More specifically, in the disk configuration formed by the merger of two neutron
stars (Sec.~\ref{ssec:test_disk_comparison}), elastic scattering causes
\begin{enumerate}
\item a decrease in average energies of neutrinos emerging from the remnant at
  all angles and for all species,
\item an increase in $\bar{\nu}_e$ and $\nu_x$ fluxes and a decrease in $\nu_e$
  fluxes viewed from along the rotation axis, and
\item a decrease in all species' fluxes viewed from the equatorial plane.
\end{enumerate}
Furthermore we find good agreement in overall number and energy luminosities
and average energies in comparisons with neutrino transport methods,
e.g.\ Monte Carlo in Sec.~\ref{ssec:test_collapse} and
M1 transport in Sec.~\ref{ssec:test_disk_comparison}.

We have also employed the ray tracing code to examine neutrino flavor
oscillations along one toy trajectory exiting the neutrino-dense environment
of the neutron star post-merger configuration (Sec.~\ref{ssec:V_nunu}).
Along that trajectory,
$25^\circ$ from the polar axis, elastic scattering increases the
$\bar{\nu}_e$ relative to the $\nu_e$ self-interaction potential,
introducing a standard matter neutrino
resonance (MNR) where there was none before.
In a similar configuration, but with matter densities lower
by a factor of a few, the effect of elastic scattering would have been to push
the MNR out to larger radii.
This example demonstrates the importance of the physics of elastic scattering
in the phenomenon of neutrino flavor oscillation.
We caution the reader, however, to avoid drawing general conclusions from this
particular example, since a single astrophysical configuration can present
dramatically different oscillation resonances along test trajectories emerging
at different angles \cite{zhu2016-mnr_nsns_remnant},
and since the MNR is extremely sensitive to a host of parameters.

Finally, we propose the following improvements to the ray tracing code to make
it a more useful and robust astrophysical simulation tool:
\begin{enumerate}
\item Use the finite-difference hydrodynamics simulation grid to represent
  background fluid and neutrino variables instead of interpolating all
  input variables to the pseudo-spectral spacetime simulation grid in order
  to begin ray tracing.
  We have found that though the interpolation of fluid and neutrino variables
  to a lower-resolution pseudo-spectral grid saves computational memory, it
  introduces more costly problems, the foremost being Gibbs-like oscillations
  at shocks and discontinuities present in fluid fields.
\item Replace explicit with implicit time stepping in the integrations along
  each ray. We have found that the stability of the time-integration demands
  extremely small step sizes of the adaptive time-stepping algorithm, especially
  for higher-energy rays. Large errors are possible if time-stepping thresholds
  are not fine-tuned to each new configuration.
\item Improve the spectral assumptions made in the \emph{gray} method.
  The test of both \emph{gray} and \emph{spectral} methods against
  the fiducial Monte Carlo calculation presented in Sec.~\ref{ssec:test_collapse}
  indicated strong agreement for overall average energies for all species. 
  However, the average energy of $\bar{\nu}_e$ emerging from the envelope
  (which contributed only a tiny fraction to the total luminosity)
  differed between the \emph{gray} treatment and the Monte Carlo by 30\%.
  Agreement could be improved with better spectral assumptions, for example
  employing pinched spectra.
\end{enumerate}
